Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-01-23 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates
update abstract

>>>>

\begin{abstract} 
We examine the flow instabilities, \ie viscous fingering on the displacement of one fluid by a second, using a novel computational method for modeling multi-phase flows in heterogeneous porous media. Those "fingers" will multiply and develop further into complex networks that will eventually effect the efficiency of field operations.

Starting from modelling the Hele-Shaw experimental apparatus, a mathematically analogous process to the two-dimensional flow in porous media problem, the finger development as a function of mobility ratio (MR), different domain set-up under uniform or different permeabilities regions and mesh types is examined. A series of cases are listed, showing the validation of the method against experimental setup followed by a more complex 2D domain setup under different viscosity ratios and different mesh resolutions. Finally, a 3D domain setup is showcasing the difference between the preferential flow path and the fingering regimes that will occur in the domain.

For the cases mentioned above, a novel control finite element method (CVFEM) is used to discretize the governing equations that will solve the multi-fluid porous media flow model. The governing multiphase porous media flow equations are solved using an unstructured mesh by introducing adaptivity in later stages. Dynamic mesh adaptivity will not only help to better capture the development of finger but also overcome the computational limitations focusing the effort where is needed, minimizing both the CPU time and the numerical diffusion.

Studying the viscous fingering patterns as a function of viscosity ratio along with the initial and boundary conditions, as well as implement adaptivity on top of an unstructured mesh will allow the user to directly apply and setup a geological model accurately and without the need of upscaling. Properties such as porosity and adaptivity can be easily introduced in the model while the idea of the preferential flow bath can be distinguished from the finger formation mechanics, in particular the tip-splitting finger behaviour.

\end{abstract}
